---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{quick_start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we're just going to show some general principles of what you need to get started in a basic pathway analysis. It acts as a quick reference to all the main points of the workflow, but more detailed examples can be found in our other tutorials

### Basic information needed for pathway analysis
The two main bits of information we need are:

- Expression tables for each condition (this can be in separate data frames, or stored within a typical Seurat object)
- Pathway gene lists

### Loading in packages 
For this simple example, we're just going to load in the SCPA package, msigdbr, and Seurat. msigdbr allows
access to a large number of gene sets that are collated [here](https://www.gsea-msigdb.org/gsea/msigdb/genesets.jsp)


```{r setup}
library(SCPA)
library(msigdbr)
library(Seurat)
```

### Generating samples from a Seurat object
Let's generate two populations to compare. To do this we'll use the `seurat_extract` function 
to extract expression matrices from our Seurat object. This function takes a Seurat object as an input,
subsets data based on the Seurat column metadata, and returns an expression file for that given subset.


```{r, eval=F}
population1 <- seurat_extract(seurat_object,
                              meta1 = "cell_type",
                              value_meta1 = "cd4_t")
population1 <- seurat_extract(seurat_object,
                              meta1 = "cell_type",
                              value_meta1 = "cd4_t")
```






