---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{quick_start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we're just going to show some general principles of what you need to get started in a basic pathway analysis. It acts as a quick reference to all the main points of the workflow, but more detailed examples can be found in our other tutorials

### Basic information needed for pathway analysis
The two main bits of information we need are:

- Expression tables for each condition (this can be in separate data frames, or stored within a typical Seurat object)
- Pathway gene lists

### Loading in packages 
For this simple example, we're just going to load in the SCPA package, msigdbr, Seurat, and dplyr. msigdbr allows access to a large number of gene sets that are collated [here](https://www.gsea-msigdb.org/gsea/msigdb/genesets.jsp)


```{r setup, results='hide', warning=F, message=F}
devtools::install_github("jackbibby1/SCPA")
library(SCPA)
library(msigdbr)
library(Seurat)
library(dplyr)
```


### Loading in data
We can use a dataset from a paper we [recently published](paper_link), which is looking
at naive CD4 T cell activation.

```{r}
load("~/Google Drive/scRNA_metabolism_paper/Processing scRNAseq data/Naive CD8/int_data_N8.RData")
naive_cd8 <- int_data_N8
```


### Generating samples from a Seurat object
Let's now generate two populations to compare. To do this we'll use the `seurat_extract` function 
to extract expression matrices from our Seurat object. This function takes a Seurat object as an input,
subsets data based on the Seurat column metadata, and returns an expression file for that given subset.


```{r}
resting <- seurat_extract(naive_cd8,
                          meta1 = "Hour",
                          value_meta1 = 0)
activated <- seurat_extract(naive_cd8,
                            meta1 = "Hour",
                            value_meta1 = 24)
```


### Generating gene sets using msigdbr
We then need to generate our gene sets. msigdbr is a handy package
that allows you to get this information. Here we're pulling all the human 'Hallmark' pathways, and
using the `format_pathways` function within SPCA to get them properly formatted. A detailed explanation of
generating gene sets for SCPA can be found [here](https://github.com/jackbibby1/SCPA/blob/main/using_gene_sets.md)

```{r}
pathways <- msigdbr("Homo sapiens", "H") %>%
    format_pathways()
```

### Comparing samples
We're all set. We now have everything that we need to compare the two populations. So just run 
`compare_pathways` and use the objects we created above.

```{r, results='hide'}
scpa_out <- compare_pathways(samples = list(resting, activated),
                             pathways = pathways)
```

And in `scpa_out`, we have all our results.

```{r}
head(scpa_out)
```


```{r, fig.width=4, fig.height=3, dpi=300, out.width="50%"}
plot_rank(scpa_out = scpa_out,
          pathway = "glycolysis",
          base_point_color = "gray70",
          highlight_point_color = "cornflowerblue")
```












