---
title: "Comparison within Seurat object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{seurat_comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We've shown that you can compare pathways if you have expression matrices, but you can also compare populations within a given Seurat object in one function within SCPA. To do this, you just need the Seurat object that should be annotated with the groups you want to compare.

Let's load in a few packages.
```{r setup, eval=F}
library(SCPA)
library(tidyverse)
library(msigdbr)
library(Seurat)
```

We're going to do the same comparison of Th1 and Tcm cells that we did in the [quick start tutorial](../docs/articles/quick_start.html), again using all Hallmark gene sets. We'll use the [Szabo, P... Sims, P](https://www.nature.com/articles/s41467-019-12464-3) T cell dataset that you can download [here](https://drive.google.com/file/d/1WSv7MLfcV_94c-C6bCUiUm_XwesCiy3N/view?usp=sharing).

```{r, eval=F}
t_cells <- readRDS("~/Google Drive/example_datasets/szabo_t_cell.rds")
```

We need to define our gene sets, and we can do this through the msigdbr package that contains a bunch of biological gene sets that can be found [here](https://www.gsea-msigdb.org/gsea/msigdb/genesets.jsp). Here we just take the Hallmark gene sets (using the "H" call to msigdbr), and use the format_pathways function to get them in a readable format for SCPA.

```{r get_pathways, eval=F}
pathways <- msigdbr("Homo sapiens", "H") %>%
  format_pathways()
```

Now to compare the Th1 and Treg populations, we can use the compare_seurat function within SCPA.

```{r compare_pathways, results='hide', eval=F}
scpa_out <- compare_seurat(t_cells,
                           group1 = "Cell", group1_population = c("Th1", "Tcm"),
                           pathways = pathways)
```

And that's it. Your results will be stored in your `scpa_out` object.





